---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tableEditor

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of tableEditor is to provide an interactive data editor.

## Notes

The app is still very much experimental and performances are not optimal.

## Installation

You can install the development version of tableEditor from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DivadNojnarg/shiny-edit-data")
```

## Example

This package allows to deploy a data editor in few steps.

1. Create a config file `config.yml`:

  - `production`: if FALSE, you are allowed to reset the data by clicking on the reset
  button (invisible in production). The default data are defined in `data_reset`.
  - `version`: Metadata displayed in the footer.
  - `board_type`: This package assumes data are stored in pins. pins can be hosted locally or on a server. Pass `local` to call `board_local` or `connect` to call `board_connect`. The latter assumes you already have defined a `CONNECT_SERVER` and `CONNECT_API_KEY` in your environment.
  - `pin_name`: Pin name within the board.
  - `filter_cols`: Leave empty to edit all columns. All passed columns will be editable.
  - `status_ok`, `status_review`, ...: The app assumes 4 data states. First, data are in `OK` state, which means they are unchanged.
  As soon as someone modifies the data, they become `IN REVIEW`. The app supports an admin mode, where one can validate or invalidate the changes resulting in 2 other data states, respectively `ACCEPTED` and `REJECTED`. Once validated any further change will reset the validation state such that data are `IN REVIEW`.

```yaml
# Below are the supported default items.
production: false
version: 0.1.0
board_type: "local"
pin_name: "<PIN_NAME>"
filter_cols: !expr c()
status_ok: "OK"
status_review: "IN REVIEW"
status_accepted: "ACCEPTED"
status_rejected: "REJECTED"
data_reset: !expr datasets::iris
```

2. Create an `app.R` script starting by the configuration linking.

```{r eval=FALSE}
options("yaml.eval.expr" = TRUE, "app.config.path" = "<PATH_TO_CONFIG>")

tableEditor::run()
```

3. Deploy to a server and enjoy.
